How to compute the fusion product of MV cycles in type A

### 2021-04-20 draft.tex
that wasnt a kilo

\section{Introduction}
% 
% The Mirkovi\'c--Vybornov isomorphism
In \cite{mirkovic2007quiver} (see also the recent sequel \cite{mirkovic2019comparison}) Mirkovi\'c and Vybornov provide a geometric version of symmetric and skew Howe $(\GL_m,\GL_n)$ dualities by relating Kazhdan--Lusztig slices in the affine Grassmannian of $\GL_m$ to slices in $n\times n$ nilpotent orbits on the one hand 
% conjugacy classes of nilpotent matrices 
and to Nakajima quiver varieties on the other hand. 
This paper generalizes the first of these relations (an isomorphism) to families of slices by way of the Beilinson--Drinfeld Grassmannian. 
Essential to \cite{mirkovic2007geometric}, this space is used to define a fusion (aka convolution) product on sheaves (representations) but can also be used to define a fusion product of MV cycles (bases). 
% 
This product structure makes the vector space spanned by the MV cycles into a commutative algebra isomorphic to the ring of functions on the unipotent subgroup (conjectured in \cite{anderson2003polytope} and proved in \cite{baumann2019Mirkovi\'c}). % conjectured by JEA and proved by BKK
% 
To quote from Mirkovi\'c and Vilonen:
\begin{quotation}
    However, we make crucial use of an idea of Drinfeld, going back to around 1990. He discovered an elegant way of obtaining the commutativity constraint by interpreting the convolution product of sheaves as a ``fusion'' product.
\end{quotation}
% 
[Liaison]

@@

We recall that these adjoint orbit closures are given by rank conditions.  More precisely we have
$$
    \overline{\OO}^\lambda = \{ A \in M_N(\CC) : \rk A^i \le N - \#\text{~boxes in first $i$ columns of }\lambda \}
$$
and
\begin{equation} 
\label{eq:ranks}
\begin{split}
    % 
    \overline{\OO}^{\lambda', \lambda''}_{0,s} = \{ A \in M_N(\CC) : &\rk A^i \le N - 
    \#\text{~boxes in first $i$ columns of }\lambda' \\
    &\rk (A-s)^i \le N - \#\text{~boxes in first $i$ columns of }\lambda'' \}
\end{split}
\end{equation}
%  
\begin{proposition} \label{prop:adjoint}
    There exists a flat family  $\overline{\OO}^{\lambda', \lambda''}_{0,\AA} \rightarrow \AA$ whose fibre over $s \in \AA$ is reduced and given by
       $ \overline{\OO}^{\lambda', \lambda''}_{0,s}$ if $s \ne 0 $ and
$        \overline{\OO}^\lambda $ if $ s = 0$.
\end{proposition}
% 
In order to prove this result, we will need to recall some results from Eisenbud--Saltman \cite{eisenbud1989rank}.
%

@@

\subsection{The Mirkovic--Vybornov slice} 
% Fix the ``$\mu$-numeration'' \((e^1_1,\ldots,e^{\mu_1}_1,\ldots,e^1_m,\ldots,e^{\mu_m}_m)\) of the standard basis of $\CC^N$. 
% \acom{I don't suppose we need this basis now that we're just describing $\TT_\mu$ in words.}
% \begin{definition}
An element of the \textbf{Mirkovi\'c--Vybornov slice} $\TT_\mu$ 
% defined as the set of $A\in M_N(\CC)$ such that 
    % \[
    %     \begin{aligned}
    %         &\text{for all } 1 \le a,s\le m\,,
    %         \text{for all } 1\le b\le \mu_a\,, 1\le t\le \mu_s\,, \\
    %         &\text{if } 1\le t < \mu_s \text{ or } t = \mu_s < b \le \mu_a \\
    %         &\text{then } e^t_s\cdot (A-J_\mu) e^b_a = 0 \,.
    %     \end{aligned}    
    % \]
% \end{definition}
% \jcom{I'm not a huge fan of this description, but I don't know how to write it better.}
% \acom{Agree. Happy to just have the in-words description.}
    % In words, $A$ 
    is a $\mu$ Jordan form matrix plus a $\mu\times\mu$ block matrix with possibly nonzero entries occuring in the first $\min(\mu_i,\mu_j)$ columns of the last row of each $\mu_i\times\mu_j$ block. 
    % $\dim \TT_\mu = N^2 - \dim \OO_\mu$. 
    In pictures, and by example, the elements $A$ for $\mu=(3,2,2,1)$ look like 
    \[
        \left[\begin{BMAT}(e){ccc;cc;c;c}{ccc;cc;c;c}
             & 1 & & & & & \\
             &  & 1 & & & & \\
            A_{11}^1 & A_{11}^2 & A_{11}^3 & A_{12}^1 & A_{12}^2 & A_{13}^1 & A_{14}^1 \\
              &  & &  & 1 & & \\
              A_{21}^1 & A_{21}^2 & & A_{22}^1 & A_{22}^2 & A_{23}^1 & A_{24}^1 \\
             A_{31}^1 & & & A_{32}^1 & & A_{33}^1 & A_{34}^1 \\
             A_{41}^1 & & & A_{42}^1 & & A_{43}^1 & A_{43}^1
        \end{BMAT}
        \right]\,. 
    \]
     To each $ A \in \TT_\mu$, we will associate the $m\times m$ matrix 
    of polynomials 
    $ g(A) = \left( g(A)_{ij} \right)$ in $ M_m(\CC[t]) $ as follows.
    \begin{equation}
        \label{eq:mvyofa}
        g(A)_{ij} = 
    \begin{cases} t^{\mu_i} - \sum_{k=1}^{\mu_i} A^k_{ji} t^{k-1} & i = j \\
            % \text{ if $ i = j$} \\
            - \sum_{k=1}^{\mu_i} A^k_{ji} t^{k-1} & i \ne j
        %  \text{ if $ i \ne j$ }
    \end{cases}
    \end{equation}
    where $A^k_{ji}$ is the $k$th entry from the left of the last row of the $\mu_j\times\mu_i$ block of $A$. 
    For example,
    \[
        \left[\begin{BMAT}(e){ccc;cc}{ccc;cc} 
            0 & 1 & 0 & 0 & 0\\
            0 & 0 & 1 & 0 & 0\\
            A_{11}^1 & A_{11}^2 & A_{11}^3 & A_{12}^1 & A_{12}^2\\
            0 & 0 & 0 & 0 & 1\\
            A_{21}^1 & A_{21}^2 & 0 & A_{22}^1 & A_{22}^2
            \end{BMAT}\right]    
            \mapsto 
            \left[\begin{array}{rr}
                t^{3} - A_{11}^3 t^2 - A_{11}^2 t - A_{11}^1 & -A_{21}^2t - A_{21}^1  \\
                -A_{12}^2 t - A_{12}^1 & t^{2} - A_{22}^2 t - A_{22}^1
            \end{array}\right]\,. 
    \]
    % For example, the $A$ above will define 
    % \acom{the transpose of! TODO. Had script?}
    % \[
    % \begin{bmatrix}
    %     t^3 - A_{11}^1 - A_{11}^2 t - A_{11}^3 t^2 & - A_{12}^1 - A_{12}^2 t & -A_{13}^1 & -A_{14}^1 \\
    %     & t^2 \\
    %     & & t \\
    %     & & & t 
    % \end{bmatrix}    
    % \]
    % \jcom{I moved the definition here since it will be useful for defining the upper triangular stuff.}
    % \acom{Ok.}
    \begin{definition} The ``\textbf{upper-triangular}'' \mvy slice $\UU^{\mu', \mu''}_{0,\AA}\rightarrow \AA $, defined by
    \UU^{\mu', \mu''}_{0,\AA} := \{ (A,s) \in \TT_\mu \times \AA : g(A)_{ii} = t^{\mu'_{i}} (t-s)^{\mu''_{i}}, g(A)_{ij} = 0 \text{ for $ j < i $ }\}\,. 
    So a matrix in $\UU^{\mu', \mu''}_{0,\AA}$ is weakly block upper-triangular and its diagonal blocks are given by the companion matrices for the polynomials $t^{\mu'_{i}} (t-s)^{\mu''_{i}}$ ($i=1,\dots,m$).
    For example, elements of $ \UU_{0,s}^{(1,1,0),(2,1,1)}$ look like 
% $\UU_{0,0}^{(?),(?)}$ entries look like 
% \[
%     \left[\begin{BMAT}(e){ccc;cc}{ccc;cc} 
%         0 & 1 & 0 & 0 & 0\\
%         0 & 0 & 1 & 0 & 0\\
%         0 & 0 & 0 & A_{12}^1 & A_{12}^2\\
%         0 & 0 & 0 & 0 & 1\\
%         0 & 0 & 0 & 0 & 0
%         \end{BMAT}\right]    
%         % \mapsto 
%         % \left[\begin{array}{rr}
%         %     t^{3} & 0 \\
%         %     -A_{12}^2 t - A_{12}^1 & t^{2}
%         % \end{array}\right]
% \]
    \[
        \left[\begin{BMAT}(e){ccc;cc;c}{ccc;cc;c} 
            0 & 1 & 0 & 0 & 0 & 0\\
            0 & 0 & 1 & 0 & 0 & 0\\
            0 & -s^2 & 2s & A_{12}^1 & A_{12}^2 & A_{13}^1 \\
            0 & 0 & 0 & 0 & 1 & 0\\
            0 & 0 & 0 & 0 & s & A_{23}^1\\
            0 & 0 & 0 & 0 & 0 & s
            \end{BMAT}\right] \,.    % \in \UU_{0,s}^{(1,1,0),(2,1,1)}
    \]
    Note that the fibre $  \UU^{\mu', \mu''}_{0,0}$ is the same as the intersection of $ \TT_\mu $ with the set of strictly upper-triangular matrices.

@@

    The the Jordan type of $A-a\big|_{E_a\cap\CC^{m(i)}}$ is given by $\lambda'(i)$ if $a = 0$ and $\lambda''(i)$ if $a = s$. 
    Moreover, the lattice $(L,s)$ defined by $g(A)$ is such that $t-a$ has the same Jordan type on 
    % $\cO_a^{m(i)} / L(a)\cap\cO_a^{m(i)} = \CC[t]^{m(i)}/L\cap\CC[t]^{m(i)}\otimes_{\CC[t]}\cO_a $ 
    \begin{equation}
        \label{eq:bcflat}
        \cO_a^{m(i)} / L(a)\cap\cO_a^{m(i)} = \CC[t]^{m(i)}/L\cap\CC[t]^{m(i)}\otimes_{\CC[t]}\cO_a 
    \end{equation}
    as does $(A-a\big|_{E_a\cap\CC^{m(i)}},s)$. 
    \acom{Not sure where to use $(?,s)$ notation and when not in this proof. Here I just didn't want to end the sentence with the bulky subscript.}
    % 
    \Cref{eq:bcflat} follows from e.g.\ \cite[Proposition~10.14]{atiyah2018introduction}:
    \begin{quotation}
        If $A$ is a Noetherian ring, $\fa$ an ideal, $\hat A$ the $\fa$-adic completion of $A$, then $\hat A$ is a flat $A$-algebra.
    \end{quotation}
    applied to $A = \CC[t]$ and $\fa = (t)$.

@@

    The second assertion is a consequence of \Cref{lem:} \acom{Cite Joel's lemma}. 
    By \Cref{adthesis} \acom{Cite my thesis} it follows that $(L(0),L(s)) \in Z'\times Z''$. 
    Since all maps are isomoprhisms the converse is automatic. 
\end{proof}

>>>>

\section{Introduction}
% 
% The Mirkovi\'c--Vybornov isomorphism
In \cite{mirkovic2007quiver} (see also the recent sequel \cite{mirkovic2019comparison}) Mirkovi\'c and Vybornov provide a geometric version of symmetric and skew Howe $(\GL_m,\GL_n)$ dualities by relating Kazhdan--Lusztig slices in the affine Grassmannian of $\GL_m$ to slices in $n\times n$ nilpotent orbits on the one hand 
% conjugacy classes of nilpotent matrices 
and to Nakajima quiver varieties on the other hand. 
This paper generalizes the first of these relations (an isomorphism) to families of slices by way of the Beilinson--Drinfeld Grassmannian. 
Essential to \cite{mirkovic2007geometric}, this space is used to define a fusion (aka convolution) product on sheaves (representations) but can also be used to define a fusion product of MV cycles (bases). 
% 
This product structure makes the vector space spanned by the MV cycles into a commutative algebra isomorphic to the ring of functions on the unipotent subgroup (conjectured in \cite{anderson2003polytope} and proved in \cite{baumann2019mirkovic}). % conjectured by JEA and proved by BKK
% 
To quote from Mirkovi\'c and Vilonen:
\begin{quotation}
    However, we make crucial use of an idea of Drinfeld, going back to around 1990. He discovered an elegant way of obtaining the commutativity constraint by interpreting the convolution product of sheaves as a ``fusion'' product.
\end{quotation}
% 
[Liaison]

@@

We recall that these adjoint orbit closures are given by rank conditions.  More precisely we have
$$
    \overline{\OO}^\lambda = \{ A \in M_N(\CC) : \rk A^i \le N - \#\text{~boxes in first $i$ columns of }\lambda \}
$$
and
\begin{equation} 
\label{eq:ranks}
\begin{split}
    % 
    \overline{\OO}^{\lambda', \lambda''}_{0,s} = \{ A \in M_N(\CC) : \rk A^i &\le N - 
    \#\text{~boxes in first $i$ columns of }\lambda' \\
    \rk (A-s)^i &\le N - \#\text{~boxes in first $i$ columns of }\lambda'' \}
\end{split}
\end{equation}
%  
\begin{proposition} \label{prop:adjoint}
    There exists a flat family  $\overline{\OO}^{\lambda', \lambda''}_{0,\AA} \rightarrow \AA$ whose fibre over $s \in \AA$ is reduced and given by
       $ \overline{\OO}^{\lambda', \lambda''}_{0,s}$ if $s \ne 0 $ and
$        \overline{\OO}^\lambda $ if $ s = 0$.
\end{proposition}
% 
In order to prove this result, we will need to recall some results from Eisenbud--Saltman \cite{eisenbud1989rank}.
%

@@

\subsection{The Mirkovic--Vybornov slice} 
% 
% Fix the ``$\mu$-numeration'' \((e^1_1,\ldots,e^{\mu_1}_1,\ldots,e^1_m,\ldots,e^{\mu_m}_m)\) of the standard basis of $\CC^N$. 
% \acom{I don't suppose we need this basis now that we're just describing $\TT_\mu$ in words.}
% 
% \begin{definition}
An element of the \textbf{Mirkovi\'c--Vybornov slice} $\TT_\mu$ 
% defined as the set of $A\in M_N(\CC)$ such that 
    % \[
    %     \begin{aligned}
    %         &\text{for all } 1 \le a,s\le m\,,
    %         \text{for all } 1\le b\le \mu_a\,, 1\le t\le \mu_s\,, \\
    %         &\text{if } 1\le t < \mu_s \text{ or } t = \mu_s < b \le \mu_a \\
    %         &\text{then } e^t_s\cdot (A-J_\mu) e^b_a = 0 \,.
    %     \end{aligned}    
    % \]
% \end{definition}
% \jcom{I'm not a huge fan of this description, but I don't know how to write it better.}
% 
% \acom{Agree. Happy to just have the in-words description.}
% 
    % In words, $A$ 
    is a $\mu$ Jordan form matrix plus a $\mu\times\mu$ block matrix with possibly nonzero entries occuring in the first $\min(\mu_i,\mu_j)$ columns of the last row of each $\mu_i\times\mu_j$ block. 
    % $\dim \TT_\mu = N^2 - \dim \OO_\mu$. 
    In pictures, and by example, the elements $A$ for $\mu=(3,2,2,1)$ look like 
    \[
        \left[\begin{BMAT}(e){ccc;cc;c;c}{ccc;cc;c;c}
             & 1 & & & & & \\
             &  & 1 & & & & \\
            A_{11}^1 & A_{11}^2 & A_{11}^3 & A_{12}^1 & A_{12}^2 & A_{13}^1 & A_{14}^1 \\
              &  & &  & 1 & & \\
              A_{21}^1 & A_{21}^2 & & A_{22}^1 & A_{22}^2 & A_{23}^1 & A_{24}^1 \\
             A_{31}^1 & & & A_{32}^1 & & A_{33}^1 & A_{34}^1 \\
             A_{41}^1 & & & A_{42}^1 & & A_{43}^1 & A_{43}^1
        \end{BMAT}
        \right]\,. 
    \]
    %

    To each $ A \in \TT_\mu$, we will associate the $m\times m$ matrix 
    of polynomials 
    % $ g(A) = \left( g(A)_{ij} \right)$ 
    $g(A)$
    in $ M_m(\CC[t]) $ whose $(i,j)$th entry is defined as follows.
    \begin{equation}
        \label{eq:mvyofa}
        g(A)_{ij} = 
    \begin{cases} t^{\mu_i} - \sum_{k=1}^{\mu_i} A^k_{ji} t^{k-1} & i = j \\
            % \text{ if $ i = j$} \\
            - \sum_{k=1}^{\mu_i} A^k_{ji} t^{k-1} & i \ne j
        %  \text{ if $ i \ne j$ }
    \end{cases}
    \end{equation}
    where $A^k_{ji}$ is the $k$th entry from the left of the last row of the $\mu_j\times\mu_i$ block of $A$. 
    For example,
    \[
        \left[\begin{BMAT}(e){ccc;cc}{ccc;cc} 
            0 & 1 & 0 & 0 & 0\\
            0 & 0 & 1 & 0 & 0\\
            A_{11}^1 & A_{11}^2 & A_{11}^3 & A_{12}^1 & A_{12}^2\\
            0 & 0 & 0 & 0 & 1\\
            A_{21}^1 & A_{21}^2 & 0 & A_{22}^1 & A_{22}^2
            \end{BMAT}\right]    
            \mapsto 
            \left[\begin{array}{rr}
                t^{3} - A_{11}^3 t^2 - A_{11}^2 t - A_{11}^1 & -A_{21}^2t - A_{21}^1  \\
                -A_{12}^2 t - A_{12}^1 & t^{2} - A_{22}^2 t - A_{22}^1
            \end{array}\right]\,. 
    \]
    % For example, the $A$ above will define 
    % \acom{the transpose of! TODO. Had script?}
    % \[
    % \begin{bmatrix}
    %     t^3 - A_{11}^1 - A_{11}^2 t - A_{11}^3 t^2 & - A_{12}^1 - A_{12}^2 t & -A_{13}^1 & -A_{14}^1 \\
    %     & t^2 \\
    %     & & t \\
    %     & & & t 
    % \end{bmatrix}    
    % \]
    % 
    % \jcom{I moved the definition here since it will be useful for defining the upper triangular stuff.}
    % \acom{Ok.}
    % 
    \acom{Can add sentence before making next definition? To the effect of ``In $\TT_\mu$ we will be interested in a certain family of block upper-triangular matrices.''}
    \begin{definition} The ``\textbf{upper-triangular}'' \mvy slice $\UU^{\mu', \mu''}_{0,\AA}\rightarrow \AA $, defined by
    $$
    \UU^{\mu', \mu''}_{0,\AA} := \{ (A,s) \in \TT_\mu \times \AA : g(A)_{ii} = t^{\mu'_{i}} (t-s)^{\mu''_{i}}, g(A)_{ij} = 0 \text{ for $ j < i $ }\}\,. 
    $$
    \end{definition}
    So a matrix in $\UU^{\mu', \mu''}_{0,\AA}$ is weakly block upper-triangular and its diagonal blocks are given by the companion matrices for the polynomials $t^{\mu'_{i}} (t-s)^{\mu''_{i}}$ ($i=1,\dots,m$).
    For example, elements of $ \UU_{0,s}^{(1,1,0),(2,1,1)}$ look like 
% 
% $\UU_{0,0}^{(?),(?)}$ entries look like 
% \[
%     \left[\begin{BMAT}(e){ccc;cc}{ccc;cc} 
%         0 & 1 & 0 & 0 & 0\\
%         0 & 0 & 1 & 0 & 0\\
%         0 & 0 & 0 & A_{12}^1 & A_{12}^2\\
%         0 & 0 & 0 & 0 & 1\\
%         0 & 0 & 0 & 0 & 0
%         \end{BMAT}\right]    
%         % \mapsto 
%         % \left[\begin{array}{rr}
%         %     t^{3} & 0 \\
%         %     -A_{12}^2 t - A_{12}^1 & t^{2}
%         % \end{array}\right]
% \]
    \[
        \left[\begin{BMAT}(e){ccc;cc;c}{ccc;cc;c} 
            0 & 1 & 0 & 0 & 0 & 0\\
            0 & 0 & 1 & 0 & 0 & 0\\
            0 & -s^2 & 2s & A_{12}^1 & A_{12}^2 & A_{13}^1 \\
            0 & 0 & 0 & 0 & 1 & 0\\
            0 & 0 & 0 & 0 & s & A_{23}^1\\
            0 & 0 & 0 & 0 & 0 & s
            \end{BMAT}\right] \,.    % \in \UU_{0,s}^{(1,1,0),(2,1,1)}
    \]
    % 
    Note that the fibre $  \UU^{\mu', \mu''}_{0,0}$ is the same as the intersection of $ \TT_\mu $ with the set of strictly upper-triangular matrices.

@@

    The the Jordan type of $A-a\big|_{E_a\cap\CC^{m(i)}}$ is given by $\lambda'(i)$ if $a = 0$ and $\lambda''(i)$ if $a = s$. 
    Moreover, the lattice $(L,s)$ defined by $g(A)$ is such that $t-a$ has the same Jordan type on 
    % $\cO_a^{m(i)} / L(a)\cap\cO_a^{m(i)} = \CC[t]^{m(i)}/L\cap\CC[t]^{m(i)}\otimes_{\CC[t]}\cO_a $ 
    \begin{equation}
        \label{eq:bcflat}
        \cO_a^{m(i)} / L(a)\cap\cO_a^{m(i)} = \CC[t]^{m(i)}/L\cap\CC[t]^{m(i)}\otimes_{\CC[t]}\cO_a 
    \end{equation}
    as does $(A-a\big|_{E_a\cap\CC^{m(i)}},s)$. 
    \acom{Not sure when to use $(?,s)$ notation and when not in this proof. Here I just didn't want to end the sentence with the bulky subscript.}
    % 
    \Cref{eq:bcflat} follows from e.g.\ \cite[Proposition~10.14]{atiyah2018introduction}:
    \begin{quotation}
        If $A$ is a Noetherian ring, $\fa$ an ideal, $\hat A$ the $\fa$-adic completion of $A$, then $\hat A$ is a flat $A$-algebra.
    \end{quotation}
    applied to $A = \CC[t]$ and $\fa = (t)$.

@@

    \acom{Here $A$ is $\CC[t]$ and $M = \cO$ right? Better, agreed. Does the direction: torsion free implies flat, or not flat implies not torsion free, hold even without the PID assumption?}

    The second assertion is a consequence of \Cref{lem:} \acom{Cite Joel's lemma}. 
    By \Cref{adthesis} \acom{Cite my thesis} it follows that $(L(0),L(s)) \in Z'\times Z''$. 
    Since all maps are isomorphisms the converse is automatic. 
\end{proof}
